A user-extensible and adaptable parser architecture

### 2008-07-30 doc/logparser-results.tex
Corrected the data used in normal-and-reverse-relative-to-shuffle graph, and added missing caption. Replaced \refwithpage with \graphref everywhere. Minor improvements.

\Graphref{Parsing time relative to shuffled ordering graph} shows the
execution time of normal and reverse orderings relative to shuffled
ordering; the mean relative execution times for different groupings of log
files are given in \tableref{Parsing time relative to shuffled ordering
table}.  Overall this optimisation provides a modest but worthwhile
performance increase of approximately 10\%, for a small investment in time
and programming.

>>>>

\Graphref{Parsing time relative to shuffled ordering graph} shows the
parsing times of normal and reverse orderings relative to shuffled
ordering; the mean relative parsing times for different groupings of log
files are given in \tableref{Parsing time relative to shuffled ordering
table}.  Overall this optimisation provides a modest but worthwhile
performance increase of approximately 10\%, for a small investment in time
and programming.

### 2008-07-30 papers/sgai-2008/paper.tex
Corrected the data used in normal-and-reverse-relative-to-shuffle graph, and added missing caption. Replaced \refwithpage with \graphref everywhere. Minor improvements.

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser author will need to write the actions required from scratch,
unless extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
Fig.~\refwithpage{Distribution of rules per action}.  The distribution is
skewed for the same reason the architecture is so successful in enabling
users to parse new log lines: most new log lines can be parsed by a
combination of a new rule and an existing action.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{REJECTION}, the action
which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_BY\_QUEUEID}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines (such as rejections), but that do not by themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects.
The parser warns about every unparsed log line, to alert the user that they
need to alter or extend their ruleset; the Postfix parser successfully
parses every input line in the \numberOFlogFILESall{} log files it is
currently being tested with.  The remaining actions have only one or two
associated rules: in some cases this is because there will only ever be one
type of log line for that action, e.g.\ all log lines showing that a remote
client has connected are matched by a single rule and handled by the
\texttt{CONNECT} action; in other cases the action is required to address a
deficiency in the log files, or a complication which arises during parsing.

@@

At the time of writing there are \numberOFrules{} different rules, with the
top 10\% matching the vast majority of the log lines, and the remaining log
lines split across the other 90\% of the rules in a Power Law distribution
(as shown in fig.~\refwithpage{rule hits graph}).
Assuming that
the distribution of log lines is reasonably steady over time, parser
efficiency should benefit from trying more frequently matching rules before
those which match less frequently.  To test this hypothesis three full test
runs were performed with different rule orderings:

@@

Figure~\refwithpage{Parsing time relative to shuffled} shows the percentage
increase of parsing times for shuffled and reversed rules.  Overall this
optimisation provides a modest but worthwhile performance increase of over
10\%, for a small investment in time and programming.
\showgraph{build/plot-hits}{Hits per rule}{rule hits graph} 
\showgraph{build/plot-normal-and-reverse-relative-to-shuffle}{Parsing time
relative to shuffled}{Parsing time relative to shuffled}

@@

The dips at log files 22 and 62--68 in fig.~\refwithpage{Parsing time
relative to shuffled} correspond to peaks in log file size in
fig.~\refwithpage{Parsing time, file size, and number of lines}, and
peaks in fig.~\refwithpage{parsing time vs file size vs number lines
factor}, where a peak means that performance is better.  The explanation
for this took some time to arrive at, but it turns out to be reasonably
simple.  The large log files in question were caused by a mail forwarding
loop, where the distribution of log lines is quite different to normal,
resulting in different performance characteristics.  The vast majority of
log lines when a mail loop occurs are from Postfix components which have a
small number of rules associated with them, whereas in general the majority
of log lines are from the Postfix component (\daemon{smtpd}) with the
highest number of rules.  Furthermore in normal log files the majority of
log lines are distributed much more evenly across rules than the majority
of log lines during a mail loop.

These two characteristics combine to reduce the average number of rules
required to parse a log line when there is a mail loop, as shown by the
peaks in fig.~\refwithpage{parsing time vs file size vs number lines
factor} and troughs in fig.~\refwithpage{Parsing time relative to
shuffled}.  When the rule ordering is reversed the majority of log lines
generated by a mail loop will be parsed using very few rules, because there
are very few rules for the Postfix components producing the log lines.
Without a mail loop the Postfix components producing the majority of log
lines have a large number of rules associated with them, and reverse
ordering deliberately starts with the least likely rules and moves towards
the most likely.  This leads to a noticeable drop in the average time
required to parse a log line during a mail loop, as shown in
fig.~\refwithpage{Parsing time relative to shuffled}.  The number of rules
which need to be consulted when the ordering is shuffled varies between the
optimal and the pessimal, with the performance varying proportionally.

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files is
\numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the total.  A
second set of statistics was generated using the minimum ruleset and the
results compared to the statistics generated using the full ruleset: the
percentage execution increase when using the full ruleset instead of the
minimal ruleset for normal, shuffled and reversed orderings is shown in
fig.~\refwithpage{Percentage execution time increase of maximum ruleset
over minimum ruleset}.  \showgraph{build/plot-normal-vs-smaller}{Percentage
execution time increase of maximum ruleset over minimum ruleset}{Percentage
execution time increase of maximum ruleset over minimum ruleset}

It is clear from fig.~\refwithpage{Percentage execution time increase of
maximum ruleset over minimum ruleset} that the increased number of rules
has a noticeable performance impact with reverse ordering, and to a lesser
extent with shuffled ordering.  The normal ordering (where are the most
frequently matching rules are tried first) shows a mean increase of 
\input{build/normal-vs-smaller-percentage-increase.tex} in
execution time for a \numberOFrulesMAXIMUMpercentage{} increase in the
number of rules.  These results show that the architecture scales very well
as the number of rules increases, and that sorting the rules is an
important optimisation enabling this.

>>>>

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser author will need to write the actions required from scratch,
unless extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{REJECTION}, the action
which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_BY\_QUEUEID}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines (such as rejections), but that do not by themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects.
The parser warns about every unparsed log line, to alert the user that they
need to alter or extend their ruleset; the Postfix parser successfully
parses every input line in the \numberOFlogFILESall{} log files it is
currently being tested with.  The remaining actions have only one or two
associated rules: in some cases this is because there will only ever be one
type of log line for that action, e.g.\ all log lines showing that a remote
client has connected are matched by a single rule and handled by the
\texttt{CONNECT} action; in other cases the action is required to address a
deficiency in the log files, or a complication which arises during parsing.

@@

At the time of writing there are \numberOFrules{} different rules, with the
top 10\% matching the vast majority of the log lines, and the remaining log
lines split across the other 90\% of the rules in a Power Law distribution
(as shown in \graphref{rule hits graph}).  Assuming that the distribution
of log lines is reasonably steady over time, parser efficiency should
benefit from trying more frequently matching rules before those which match
less frequently.  To test this hypothesis three full test runs were
performed with different rule orderings:

@@

\Graphref{Parsing time relative to shuffled} shows the parsing times of
normal and reversed orderings relative to shuffled ordering.  Overall this
optimisation provides a modest but worthwhile reduction in parsing time of
approximately 10\%.  \showgraph{build/plot-hits}{Hits per rule}{rule hits
graph}
\showgraph{build/plot-normal-and-reverse-relative-to-shuffle}{Parsing time
relative to shuffled ordering}{Parsing time relative to shuffled}

@@

The dips at log files 22 and 62--68 in \graphref{Parsing time relative to
shuffled} correspond to peaks in log file size in \graphref{Parsing time,
file size, and number of lines}, and peaks in \graphref{parsing time vs
file size vs number lines factor}, where a peak means that performance is
better.  The explanation for this took some time to arrive at, but it turns
out to be reasonably simple.  The large log files in question were caused
by a mail forwarding loop, where the distribution of log lines is quite
different to normal, resulting in different performance characteristics.
The vast majority of log lines when a mail loop occurs are from Postfix
components which have a small number of rules associated with them, whereas
in general the majority of log lines are from the Postfix component
(\daemon{smtpd}) with the highest number of rules.  Furthermore in normal
log files the majority of log lines are distributed much more evenly across
rules than the majority of log lines during a mail loop.

These two characteristics combine to reduce the average number of rules
required to parse a log line when there is a mail loop, as shown by the
peaks in \graphref{parsing time vs file size vs number lines factor} and
troughs in \graphref{Parsing time relative to shuffled}.  When the rule
ordering is reversed the majority of log lines generated by a mail loop
will be parsed using very few rules, because there are very few rules for
the Postfix components producing the log lines.  Without a mail loop the
Postfix components producing the majority of log lines have a large number
of rules associated with them, and reverse ordering deliberately starts
with the least likely rules and moves towards the most likely.  This leads
to a noticeable drop in the average time required to parse a log line
during a mail loop, as shown in \graphref{Parsing time relative to
shuffled}.  The number of rules which need to be consulted when the
ordering is shuffled varies between the optimal and the pessimal, with the
performance varying proportionally.

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files is
\numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the total.  A
second set of statistics was generated using the minimum ruleset and the
results compared to the statistics generated using the full ruleset: the
percentage execution increase when using the full ruleset instead of the
minimal ruleset for normal, shuffled and reversed orderings is shown in
\graphref{Percentage execution time increase of maximum ruleset over
minimum ruleset}.  \showgraph{build/plot-normal-vs-smaller}{Percentage
execution time increase of maximum ruleset over minimum ruleset}{Percentage
execution time increase of maximum ruleset over minimum ruleset}

It is clear from \graphref{Percentage execution time increase of maximum
ruleset over minimum ruleset} that the increased number of rules has a
noticeable performance impact with reverse ordering, and to a lesser extent
with shuffled ordering.  The normal ordering (where are the most frequently
matching rules are tried first) shows a mean increase of
\input{build/normal-vs-smaller-percentage-increase.tex} in execution time
for a \numberOFrulesMAXIMUMpercentage{} increase in the number of rules.
These results show that the architecture scales very well as the number of
rules increases, and that sorting the rules is an important optimisation
enabling this.
